---
title: "Moving Data in htn"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
##Introduction
If you've completed the [htn tutorial]('Getting_started.html'), you've created a network, graphed it, and found a community that you are interested in exploring further.  But how do you get the data out of the graph?  Which documents are part of your chosen community?

In this tutorial, we'll cover adding and extracting data from a `docNetwork` object so that you can do more with your graphs.  To begin, we'll create our `docNetwork` object and graph it (in case you haven't done the [htn tutorial]('Getting_started.html')).

```{r, fig.height=8, fig.width=8, warning=F}
library(tei2r)
library(htn)
library(igraph)
dl = buildDocList(directory = system.file('extdata/natlaw/', package='tei2r'), indexFile=system.file('extdata/natlaw/index.csv', package='tei2r'))
dnet = buildNetwork(dl, view=T)
drawGraph(dnet@graph, comms=dnet@communities)
```

Now let's find out which documents are in our chosen community.  We can extract the `TCP` numbers for the documents pretty easily by running:
```{r, warning=F}
results = extractData(dnet = dnet, comms=dnet@communities, file="", sub=T, by=2)
results
```